---
title: "vignette for rdtest: A joint diagnosis for regression discontinuity designs."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*rdtest* offers a joint diagnostic procedure for the testable restrictions of
regression discontinuity (RD) designs. This package combines the existing
packages of diagnostic testing (rddensity and rdrobust) to offer a set of joint
test statistics to evaluate the validity of a given RD design. The theoretical
background of this package is provided in [Fusejima, Ishihara, and Sawada (2023,
arXiv)](https://arxiv.org/abs/2205.04345).

## Background
The sufficient condition for identification in RD design, local randomization,
has two implications: continuous density function and balancing of
pre-treatment covariates. Violation of any of these testable restrictions raises
a concern about the sufficient condition. (Nevertheless, these testable
restrictions are not directly related to identification itself. See [Ishihara
and Sawada (2023)](https://arxiv.org/abs/2009.07551) for further details.)

Both testable restrictions are implemented in R packages, for example,
[*rddensity*](https://rdpackages.github.io/rddensity/) ([Calonico, Cattaneo, and Titiunik (2014)](https://onlinelibrary.wiley.com/doi/10.3982/ECTA11757)) for the former and [*rdrobust*](https://rdpackages.github.io/rdrobust/) ([Cattaneo, Jansson, and Ma (2020)](https://www.tandfonline.com/doi/abs/10.1080/01621459.2019.1635480)) for the latter. As pointed
out in [Fusejima, et al. (2023)](https://arxiv.org/abs/2205.04345), however,
multiple tests (median number of tests is 12) are run in practice without
appropriate size control for most of the studies. Consequently, these RD
practices suffer from massive size distortion due to *multiple testing problem*.
Among 60 papers studied, 35\% of studies reject at least one of these testable
restrictions; however, the rejection of each test separately is approximately 5\%.
Hence, the RD designs in these studies appear to maintain the null hypotheses;
however, not a small fraction of studies contain false rejections of the null
hypotheses more frequently than they should be. Our *rdtest* package is the
resolution to this multiple testing problem.

## Installation

You can install the development version of rdtest from [GitHub](https://github.com/) with:
install.packages("devtools")
devtools::install_github("SMasa11/rdtest")

## Usages

`Z` may be either a numeric vector or a data.frame containing covariates.
```{r setup}
library(rdtest)
library(haven)
set.seed(1)
N <- 1000
Z <- data.frame(var1 = rnorm(N),
                var2 = rnorm(N),
                var3 = rnorm(N),
                var4 = rnorm(N),
                var5 = rnorm(N))
vec_X <- rnorm(N)
```

Default option is standardized Wald statistics (`sWald`).
```{r basic usage}
res_rd <- rdtest::rdtest(Z = Z, vec_X = vec_X)
ans <- summary(res_rd)
```

By toggling `bool_L2_std` option FALSE, `rdtest` returns non-standardized Wald test result, but in simulation, the non-standardized test does not control its size appropriately.
```{r Not recommended}
res_rd <- rdtest::rdtest(Z = Z, vec_X = vec_X, bool_L2_std = FALSE)
ans <- summary(res_rd)
```


With `bool_max_test` option TRUE, `rdtest` returns max test result instead of `sWald` test result.
```{r other statistics}
res_rd <- rdtest::rdtest(Z = Z, vec_X = vec_X, bool_max_test = TRUE, bool_equivalence = TRUE)
ans <- summary(res_rd)
```
When `bool_max_test` is TRUE, `bool_equivalence` option is available.
If `bool_equivalence` is TRUE, the equivalence test result will also be returned.
It returns the maximum (in absolute value) $\epsilon$ value for which we can reject the null hypothesis of the max of test stastitics being larger than $\epsilon$ in absolute value against the alternative of the test statistic being smaller than $\epsilon$ in absolute value.

## Use cases:
### Clark and Royer (2013, AER)
Clark and Royer (2013) study the impact of the compulsory schooling changes in 
UK on log odds of adulthood death rates. The policy changes prohibit students 
from leaving school before grade nine (the 1947 reform) and from leaving school 
before grade ten (the 1972 reform).

In their Appendix Figure A1, they check the absence of discontinuity in births 
and stillborn among the cohorts around the policy changes in 1947 and 1972 to
verify that the differences in adulthood deaths are not driven by the 
differences in the number of births and stillborn children.

We confirm that their results are robust as the joint null hypothesis of no 
difference whatsoever for the 1947 policy change cohorts
```{r 1947 change}
data_path <- system.file("extdata", "018_births.dta", package = "rdtest")
df_018 <- haven::read_stata(data_path)
df_018_1 <- data.frame(df_018[df_018$r == 1,c("births","still","s1mob1")])
res_rd <- rdtest::rdtest(Z=df_018_1[,c("births","still")],vec_X = df_018_1$s1mob1)
ans <- summary(res_rd)

res_rd <- rdtest::rdtest(Z=df_018_1[,c("births","still")],vec_X = df_018_1$s1mob1,bool_max_test = TRUE,bool_equivalence = TRUE)
ans <- summary(res_rd)
```
as well as the 1972 policy change cohorts.
```{r 1972 change}
df_018 <- haven::read_stata(data_path)
df_018_2 <- data.frame(df_018[df_018$r == 2,c("births","still","s2mob1")])
vec_nan <- is.nan(df_018_2[,1]) + is.na(df_018_2[,1])
vec_nan <- vec_nan + is.nan(df_018_2[,2]) + is.na(df_018_2[,2])
vec_nan <- vec_nan + is.nan(df_018_2[,3]) + is.na(df_018_2[,3])
df_018_2 <- df_018_2[vec_nan == 0,]

res_rd <- rdtest::rdtest(Z=df_018_2[,c("births","still")],vec_X = df_018_2$s2mob1)
ans <- summary(res_rd)

res_rd <- rdtest::rdtest(Z=df_018_2[,c("births","still")],vec_X = df_018_2$s2mob1,bool_max_test = TRUE,bool_equivalence = TRUE)
ans <- summary(res_rd)
```
Note that the equivalence test shows the null hypothesis that the max of test statistics is larger than an almost zero $\epsilon$ in absolute value can be rejected. This result comes from the fact that the density test passes with the perfect balance of zero t-statistics.

### Pons and Tricaud (2018, Ecta)
Pons and Tricaud (2018) study the impact of qualifying the third candidate for
the second round of French parliamentary and local elections on the share of 
registered citizens for the second round and shares of the top two candidates.

French parliamentary and local elections are in two rounds. The top two 
candidates of the first round are always qualified for the second round, but the
third candidate is not necessarily. For the third candidates being qualified,
their first-round vote share must be higher than 12.5 percent. The authors 
exploit the threshold of having the third candidate as a fuzzy design on the 
voter participation rates.

In their Appendix Table B-I, they conduct placebo tests for five different 
pre-determined covariates: the number of registered citizens, number of 
candidates in the first round, turnout rate of the first round, the distance of 
the top-two vote share in the first round, and the predicted assignment 
probability of actually having the third candidates (not their qualification).

We redo their analysis in our joint test for the joint null hypothesis of no 
differences in any of the covariate means and density whatsoever. We confirm
that the joint null hypothesis is not rejected, supporting the validity of the 
original analysis further.
```{r}
data_path <- system.file("extdata", "165_analysis.dta", package = "rdtest")
df_165 <- haven::read_stata(data_path)
list_var <- c("nb_registered_R1",
              "nb_candidates_R1",
              "prop_registered_turnout_R1",
              "distance_voteshare_cand12_R1",
              "predicted_assignment")
df_165 <- data.frame(df_165[,c(list_var,"running")])
vec_nan <- is.nan(df_165[,1]) + is.na(df_165[,1])
for (i in seq(2,6)) {
  vec_nan <- vec_nan + is.nan(df_165[,i]) + is.na(df_165[,i])
}
print(sum((vec_nan > 0)))
df_165 <- df_165[vec_nan == 0,]

res_rd <- rdtest::rdtest(Z=df_165[,list_var],vec_X = df_165$running)
ans <- summary(res_rd)

res_rd <- rdtest::rdtest(Z=df_165[,list_var],vec_X = df_165$running,bool_max_test = TRUE,bool_equivalence = TRUE)
ans <- summary(res_rd)
```

### References for the applciations

Clark, Damon, and Heather Royer. “The Effect of Education on Adult Mortality and Health: Evidence from Britain.” American Economic Review 103, no. 6 (October 1, 2013): 2087–2120. https://doi.org/10.1257/aer.103.6.2087.

Data citation: Clark, Damon, and Royer, Heather. Replication data for: The Effect of Education on Adult Mortality and Health: Evidence from Britain. Nashville, TN: American Economic Association [publisher], 2013. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2019-10-11. https://doi.org/10.3886/E112669V1

Pons, Vincent, and Clémence Tricaud. “Expressive Voting and Its Cost: Evidence From Runoffs With Two or Three Candidates.” Econometrica 86, no. 5 (2018): 1621–49. https://doi.org/10.3982/ECTA15373.

