---
title: "vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`Z` may be either a numeric vector or a data.frame containing covariates.
```{r setup}
library(rdtest)
set.seed(1)
N <- 1000
Z <- data.frame(var1 = rnorm(N),
                var2 = rnorm(N),
                var3 = rnorm(N),
                var4 = rnorm(N),
                var5 = rnorm(N))
vec_X <- rnorm(N)
```

Default option is standardized Wald statistics (`sWald`).
```{r basic usage}
res_rd <- rdtest::rdtest(Z = Z, vec_X = vec_X)
ans <- summary(res_rd)
```

With `bool_max_test` option TRUE, `rdtest` returns max test result instead of `sWald` test result.
```{r other statistics}
res_rd <- rdtest::rdtest(Z = Z, vec_X = vec_X, bool_max_test = TRUE)
ans <- summary(res_rd)
```

By toggling `bool_L2_std` option FALSE, `rdtest` returns non-standardized Wald test result, but in simulation, the non-standardized test does not control its size appropriately.
```{r Not recommended}
res_rd <- rdtest::rdtest(Z = Z, vec_X = vec_X, bool_L2_std = FALSE)
ans <- summary(res_rd)
```
